### 2021-04-15 newvariant.tex
minor changes

Instead of estimating relative strength from speed, one can estimate time-varying or instantaneous reproduction numbers $\RR(t)$ of the variant and the wild type from incidence data \citep{fraser2007estimating}, and directly compare their ratios;
such methods have been used in previous analyses of the B.1.1.7 strain by \cite{volz2021transmission}.
Assuming that the generation-interval distribution remains constant, the instantaneous reproduction numbers of the new variant and of the wild type can be estimated from their corresponding incidence curves---an approach popularized by \cite{cori2013new}:
\begin{equation}
\RR_x(t) = \frac{i_x(t)}{\int_0^\infty i_x(t-\tau) g_x(\tau) \dtau}.
\end{equation}
Under strength-like intervention measures that reduce transmission rates by a constant amount, we expect ratios between reproduction numbers to remain constant and correspond to the true relative strength: $\Rv(t)/\Rw(t) = \rho$.
However, if the assumed generation-interval distribution $\hat{g}(\tau)$ is different from the true distribution, then the ratio between the estimated reproduction numbers $\hat{\rho}(t) = \hat{\RR}_{\textrm{var}}(t)/\hat{\RR}_{\textrm{wt}}(t)$ may change, even if the true ratio does not.

@@

However, a fixed generation generation interval assumed by \cite{davies2021estimated} is too narrow.
The mean generation interval of 6.5 days assumed by \cite{volz2021transmission} is also likely too long:
first, this estimate actually corresponds to the observed mean serial intervals from China in between January and February, 2020 \citep{bi2020epidemiology}, during which they were more likely to observe longer serial intervals due to initial epidemic growth \citep{park2021forward, ali2020serial}; and second, recently infected individuals may be less likely to transmit after symptom onset due to increased awareness of the pandemic compared to earlier on, leading to shorter generation intervals.
Instead, if we replace the assumed generation-interval distributions for \cite{volz2021transmission} with shorter generation interval estimates from Tianjin, China ($\Gw=2.57\days$ and $\kappa=1$; \cite{ganyani2020estimating}), we obtain $\rho=1.27$, which is consistent with the attack rate analysis \citep{ukinvest}.
We do not claim that these estimates of $\rho$ or $\Gw$ are more accurate;
instead, this back-of-the-envelope calculation highlights the importance of propagating uncertainty in the generation-interval distribution in assessing the relative strength and speed of SARS-CoV-2 variants.

@@

We use a two-strain renewal equation to simulate the spread of the variant and the wild type.
Ignoring birth and death, the incidence of infection caused by the variant $\iv$ and the wild type $\iw$ assuming perfect cross immunity is given by:
\begin{align}
\frac{dS}{dt} &= - \iv(t) - \iw(t)\\
\iv(t) &= \Rv(t) \int_{0}^\infty \iv(t-\tau) g_{\mathrm{var}}(\tau) \dtau\\
\iw(t) &= \Rw(t) \int_{0}^\infty \iw(t-\tau) g_{\mathrm{wt}}(\tau) \dtau
\end{align}
where $S$ represents the proportion of susceptible individuals.
We discretize the model at the time scale of 0.025 days as described in \citep{park2021forward} and simulated the model with $\iv(0) = 0.001$ and $\iw(0) = 0.1$.
We assume that $\Rw(t)$ remains at 2 for 30 days, immediately decreases and stays at 0.5 until day 60, and immediately increases back to 1.2.
Finally, we assume $\Rv(t) = \rho \Rw(t)$.

>>>>

Instead of estimating relative strength from speed, one can estimate time-varying or instantaneous reproduction numbers $\RR(t)$ of the variant and the wild type from incidence data \citep{fraser2007estimating}, and directly compare their ratios;
such methods have been used in previous analyses of the B.1.1.7 strain by \cite{volz2021transmission}.
Assuming that the generation-interval distribution remains constant, the instantaneous reproduction numbers of the new variant and of the wild type can be estimated from their corresponding incidence curves---an approach popularized by \cite{cori2013new}:
\begin{equation}
\RR_x(t) = \frac{i_x(t)}{\int_0^\infty i_x(t-\tau) g_x(\tau) \dtau}.
\label{eq:rt}
\end{equation}
Under strength-like intervention measures that reduce transmission rates by a constant amount, we expect ratios between reproduction numbers to remain constant and correspond to the true relative strength: $\Rv(t)/\Rw(t) = \rho$.
However, if the assumed generation-interval distribution $\hat{g}(\tau)$ is different from the true distribution, then the ratio between the estimated reproduction numbers $\hat{\rho}(t) = \hat{\RR}_{\textrm{var}}(t)/\hat{\RR}_{\textrm{wt}}(t)$ may change, even if the true ratio does not.

@@

However, a fixed generation generation interval assumed by \cite{davies2021estimated} is too narrow.
The mean generation interval of 6.5 days assumed by \cite{volz2021transmission} is also likely too long:
first, this estimate actually corresponds to the observed mean serial intervals from China in between January and February, 2020 \citep{bi2020epidemiology}, during which they were more likely to observe longer serial intervals due to initial epidemic growth \citep{park2021forward, ali2020serial}; and second, recently infected individuals may be less likely to transmit after symptom onset due to increased awareness of the pandemic compared to earlier on, leading to shorter generation intervals.
Instead, if we replace the assumed generation-interval distributions for \cite{volz2021transmission} with shorter generation interval estimates from Tianjin, China ($\Gw=2.57\days$ and $\kappa=1$; \cite{ganyani2020estimating}), we obtain $\rho=1.27$, which is consistent with the attack rate analysis \citep{ukinvest}.
We do not claim that these estimates of $\rho$ or $\Gw$ are more accurate;
instead, this back-of-the-envelope calculation highlights the importance of propagating uncertainty in the generation-interval distribution in assessing the relative strength and speed of SARS-CoV-2 variants.
Obtaining an accurate and precise estimate of $\rho$ will need to take into account changes in $\rho$ as well as underlying $\rw$.

@@

We use a two-strain renewal equation to simulate the spread of the variant and the wild type.
Ignoring birth and death, the incidence of infection caused by the variant $\iv$ and the wild type $\iw$ assuming perfect cross immunity is given by:
\begin{align}
\frac{dS}{dt} &= - \iv(t) - \iw(t)\\
\iv(t) &= \Rv(t) \int_{0}^\infty \iv(t-\tau) g_{\mathrm{var}}(\tau) \dtau\\
\iw(t) &= \Rw(t) \int_{0}^\infty \iw(t-\tau) g_{\mathrm{wt}}(\tau) \dtau
\end{align}
where $S$ represents the proportion of susceptible individuals.
We discretize the model at the time scale of 0.025 days as described in \citep{park2021forward} and simulated the model with $\iv(0) = 0.001$ and $\iw(0) = 0.1$.
We assume that $\Rw(t)$ remains at 2 for 30 days, immediately decreases and stays at 0.5 until day 60, and immediately increases back to 1.2.
Finally, we assume $\Rv(t) = \rho \Rw(t)$.
From simulated incidence curves between day 15 and 70, we estimate the instaneous reproduction number using \eref{rt}.
We ignore incidence before day 15 to remove any potential transient effects.
